---
title: "CloudOS User Training - Interactive Analysis (RStudio)"
output: html_document
date: "2023-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The structure of this notebook follows the structure of the training session:
For every workflow or set of tasks, the trainers will first walk you through it, and show you how it s carried out Step-by-Step. The code for the step-by-step explanation of every workflow is available in corresponding step-by-step sections of this notebook.

## Work interactively in RStudio - I
### Step-by-step
- Create a bew Rstudio session, loading in your cohort 
- Install and load libraries
- Read in our dataset
- Run an analysis over the data and create a plot

```{r cars}
install.packages("survival")
install.packages("ggplot2")
install.packages("ggfortify")
install.packages("plotly")
install.packages("dplyr")
install.packages("tidyverse")
install.packages("DT")
install.packages("shiny")

library(survival)
library(ggplot2)
library(ggfortify)
library(plotly)
library(dplyr)
library(tidyverse)
library(DT)
library(shiny)
```

```{r}
data <- read.csv("synthetic_dataset.csv", header=TRUE)
```


```{r}
head(data)
```

```{r}
# Summary Statistics 
summary(data[c("height", "weight", "BMI" )]) 

# Bar plot for categorical variables 

barplot(table(data$sex), main = "Distribution by Sex" )
barplot(table(data$ethnicity), main = "Distribution by Sex")
barplot(table(data$vitalstatus), main = "Distribution by Vital Status")
barplot(table(data$cancer_disease_type), main = "Distribution by Cancer Disease Type")
barplot(table(data$smoking_status), main = "Distribution by Smoking Status")

```
#Exercises 

1. Create a binary classification model using logistic regression to precict vital status based on: 
- sex 
- ethnicityname 
- cancer_disease_type
- smoking_status

Before fitting the model, convert vitalstatus to a binary variable 

```{r, echo=FALSE}
# Solution 1 

#Convert vitalstatus to a binary variable 
data$vital_status_alive <- ifelse(data$vitalstatus == "A", 1, 0)

#Fitting the model, printing the summary as an output 
model <- glm(vital_status_alive ~ sex + ethnicityname + cancer_disease_type + smoking_status, data=data, family = binomial)
summary(model)

```

2. Create a regression model to predict BMI based on: 
- height
- weight 
- smoking_status
- alcohol_consumption 

```{r}
# Regression model: predicting BMI 

model <- lm(BMI ~ height + weight + smoking_status + alcohol_consumption, data = data)

summary(model)

```

3. Create an interactive plot of height vs weight, with the points colored by cancer_disease_sub_type 

```{r}
# Interactive scatter plot with plotly 

# Load the magrittr package
library(magrittr)


plot_ly(data = data, x = ~height, y = ~weight, color= ~cancer_disease_sub_type,
        type = "scatter", mode = "markers", marker = list(size = 10)) %>%
  layout(xaxis = list(title = "Height"), yaxis = list(title = "Weight"))

```
4. Run this kernel to see how you can create a mini app with your plot and interactive output 

```{r}
#Interactive filtering with Shiny 


ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput("disease_type", "Cancer Disease Type", choices = unique(data$cancer_disease_type))
    )
    , 
    mainPanel(
      plotlyOutput("scatter_plot")
    )
  ))

  server <- function(input, output) {
    output$scatter_plot <- renderPlotly({
      filtered_data <- subset(data, cancer_disease_type == input$disease_type)

      plot <- plot_ly(data = filtered_data, x = ~height, y= ~weight, color = ~cancer_disease_sub_type,
        type = "scatter", mode = "markers", marker = list(size = 10)) %>% 
        layout(xaxis = list(title = "Height"), yaxis = list(title = "Weight"))
    })
  }
  
  shinyApp(ui,server)

``` 